---
title: "Tidytuesday - Spotify Song"
author: "Amri Rohman"
date: "21/5/2020"
output: 
  html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width=10,
                      fig.height=3.5)
options(scipen = 999,
        digits = 2) 
```
![](https://images.unsplash.com/photo-1504509546545-e000b4a62425?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1350&q=80)

# About Dataset
***

The data this week comes from Spotify via the [`spotifyr` package](https://www.rcharlie.com/spotifyr/). [Charlie Thompson](https://twitter.com/_RCharlie), [Josiah Parry](https://twitter.com/JosiahParry), Donal Phipps, and Tom Wolff authored this package to make it easier to get either your own data or general metadata arounds songs from Spotify's API. Make sure to check out the [`spotifyr` package](https://www.rcharlie.com/spotifyr/) website to see how you can collect your own data!

[Kaylin Pavlik](https://twitter.com/kaylinquest/status/1213138536570015745) had a recent [blogpost](https://www.kaylinpavlik.com/classifying-songs-genres/) using the audio features to explore and classify songs. She used the `spotifyr` package to collect about 5000 songs from 6 main categories (EDM, Latin, Pop, R&B, Rap, & Rock). 

[More Information About Dataset.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md)

# Loading Library and Data
***

```{r}
library(tidyverse)
library(janitor)
library(knitr)
library(skimr)

spotify_songs_raw <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```
I used several packages to help me do this task. That is Tidyverse, janitor, skimr and knitr

# Data Preparation
***

```{r}
spotify_songs <- spotify_songs_raw %>% 
  mutate(
    mode = case_when(mode == 1 ~ "major",
                     mode == 0 ~ "minor"),
    track_album_release_date = as.Date(track_album_release_date),
    year = as.numeric(format(track_album_release_date,'%Y')),
    era = case_when(year <= 2000 ~ "under 00s era",
                    year > 2000 & year <= 2010 ~ "2010 era",
                    year > 2010 ~ "modern era"),
    duration_min = duration_ms/60000) %>%
  select(-track_id, -track_album_id, -playlist_id, -duration_ms)

spotify_songs %>% 
  select(track_name, track_artist, duration_min,
         track_popularity, playlist_genre, 
         tempo) %>%
  head() %>%
  kable()
```

# Summary
***
### Numeric Variable

```{r}
spotify_songs %>% 
  skim() %>%
  yank('numeric') %>% 
  select(-p0, -p25, -p75, -p100) %>%
  kable() 
```

### Character Variable

```{r}
spotify_songs %>% 
  skim() %>%
  yank('character') %>% 
  select(-whitespace, -empty) %>%
  kable() 
```

### Date Variable

```{r}
spotify_songs %>% 
  skim() %>%
  yank('Date') %>% 
  #select() %>%
  kable() 
```

# Data Exploration
***
### What is the most popular song?

```{r}
spotify_songs %>%
  select(track_name, track_artist, track_popularity) %>%
  arrange(-track_popularity) %>%
  distinct() %>%
  filter(track_popularity %in% c(97:100)) %>% 
  kable()
```

The first most popular song in Spotify is Dance Monkey - Tones and I. Followed by ROXANNEArizona Zervas and the others.

### How is the dominance of mode/modality in every genre?

```{r}
spotify_songs %>%
  group_by(mode, playlist_genre) %>%
  count(mode, sort = TRUE) %>%
  ggplot() +
  geom_bar(aes(mode, n, fill=mode), stat="identity") +
  labs(x='Proportion', y='Playlist Genre') +
  facet_wrap(~playlist_genre, nrow = 2)+
  theme_minimal() +
  theme(legend.position = 'bottom')
```
Major mode/modality is most often used in all genres on Spotify.

### How is the dominance of mode/modality in every subgenre?

```{r fig.height=4}
spotify_songs %>%
  group_by(playlist_genre, mode) %>%
  count(playlist_subgenre) %>%
  ggplot() +
  geom_bar(aes(reorder(playlist_subgenre, n), n, fill=mode), position = 'stack', stat='identity') +
  coord_flip() +
  facet_wrap(playlist_genre~., scales="free", nrow=3) +
  labs(x='Sub Genre', y='Count') +
  theme_minimal() +
  theme(legend.position = 'bottom')
```
The sub genre of rock has the most dominant major mode/modality than other sub genres. The sub genre of rap has a balanced mode/modality relatively.

### Which song from the era is the most on Spotify?

```{r}
spotify_songs %>%
  group_by(era) %>%
  count(era, sort = TRUE) %>%
  na.omit() %>%
  ggplot(aes(era, y=n, fill=era)) +
  geom_bar(stat="identity") +
  labs(x='Era', y='Count') +
  theme_minimal()
```
song from the modern era (released above in 2010) most in spotify.

### How about the tempo in each genre and Era?

```{r}
spotify_songs %>%
  group_by(playlist_genre, era) %>%
  na.omit() %>%
  ggplot() +
  geom_boxplot(aes(era, tempo, fill=era), outlier.size=0) +
  facet_grid(~playlist_genre) + 
  coord_flip() +
  labs(x='Tempo') +
  theme_minimal() +
  theme(legend.position = 'none')
```
Rap song in the under 00s era had a slower tempo compared to other eras. EDM song has small variation about tempo in every era.

### How the distribution of duration in each Genre and Era?

```{r}
spotify_songs %>%
  na.omit(era) %>%
  ggplot() +
  geom_density(aes(duration_min, fill=playlist_genre)) +
  facet_grid(era~playlist_genre) +
  theme_minimal() +
  labs(x='Duration (min)') +
  theme(legend.position = 'none')
```
EDM song in the 2010 era has a relatively longer duration than in other eras. Rap song in the modern era has a relatively shorter duration than in other eras. While others have relatively the same duration in each era.

### How about audio Features in every genre?

```{r}
spotify_songs %>%
    select(danceability, energy, -loudness, speechiness, acousticness, instrumentalness, liveness, valence,
           playlist_genre,
           -track_popularity, -key, -tempo, -year, -duration_min) %>% 
  gather(key=status, value=value, -playlist_genre) %>%
  ggplot() +
  geom_bar(aes(status, value, fill=status), stat='identity') +
  coord_flip() +
  facet_wrap(~playlist_genre, nrow=2) +
  theme_minimal() +
  labs(x='Value') +
  theme(legend.position = 'none')
```



# Thank You
Amri Rohman. <br>
Sidoarjo, East Java, ID
